### 2023-04-11 paper/paper.tex
Got the references correct. I need another 4 pages
content and some pretty diagrams.

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{tabularx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Towards automated provenance collection for experimental runs of agent-based models\thanks{Supported by The Scottish Government}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Doug Salt\inst{1}\orcidID{0000-1111-2222-3333} \and
Gary Polhill\inst{2,3}\orcidID{1111-2222-3333-4444} }
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{The James Hutton Institute, Craigiebuckler Aberdeen AB15 8QH Scotland
\email{doug.salt@hutton.ac.uk}\\
\url{http://hutton.ac.uk}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The abstract should briefly summarize the contents of the paper in
15--250 words.

\keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}
%
%
%
\section{Introduction}

Replication of social simulation results has been highlighted as a significant
issue for the ABM for a number of years (e.g.  \cite{edmonds2003replication}),
and in particular the paper that forms the basis of this work
\cite{polhill2017lessons}.  The focus of replication work has previously just
been on the model itself, but as was shown in ibid, the analysis of the
outputs of the model can potentially be just as complex, and no less difficult
to replicate unless adequate records are kept. Indeed there as an additional
attempt to reproduce the results and even with the lessons learnt from ibid, it
still took months to recreate the final diagrams that were submitted to paper.
The TRACE protocol \cite{schmolke2010ecological,ayllon2021keeping} provides
some guidance highlighting the need to keep a notebook of the analysis done and
a standardised approach to making that notebook; however, the lessons learned
from the replication exercise in \cite{polhill2017lessons} show that more
detailed guidance on the information that should be recorded,  and on tools
that could be used to support the process. Ideally the aim should be to
completely automate the process, record accurately sources of data used,
version check applications used essentially providing a complete graph from
data to result.

The output analysis replication in this paper concerns earlier work with
FEARLUS-SPOMM, which is a coupled agent-based model of agricultural decision-
making and species stochastic patch occupancy metacommunity model that has been
used to explore incentivisation strategies to improve biodiversity
\cite{polhill_nonlinearities_2013}, \cite{gimona2011exploring}. Belonging to the
‘typification’ class of social simulations  \cite{boero2005does}.  This work
involved the analysis of the outputs of approximately 20,000 runs of the model
using a number of techniques aimed at demonstrating nonlinearities in the
relationship between incentivisation and biodiversity outcome.  Recording
workflow data on the process used to create analysis can be challenging, and
currently there are no codified standards as to how this should be done for
ABMs. For FEARLUS-SPOMM, the methods used drew heavily on statistical
techniques available as R packages that are as part of core R functionality.
Although R allows transcripts of interactive terminal sessions to be saved, the
work involved great deal of exploration of different ways of attempting to
visualise and analyse the nonlinearities in the model results, not all of
which were likely to be reported in the paper. Such logs are therefore not the
best way to record the means by which the outputs were analysed, and hence the
strategy used was to save each analysis or visualisation in a(n R) script.
Since the output from the (Swarm) software that generated the output data being
analysed used a mixture of text formats, some Perl scripts were also written to
process that output into a CSV file for easy import into R. When the MIRACLE pr
ject \cite{parker2019final} provided a context in which the replication of that
analysis was necessary, an opportunity was created to test the viability of the
above strategy.

In the rest of this paper, we describe  a tool for automatically recording
metadata, which can be incorporated into the analysis replication process, and
how this  was used to regenerate some of the figures in
\cite{polhill_nonlinearities_2013} and we test a tool to support keeping the
records needed to per- form replication more easily, before making some
recommendations in concluding remarks.

\section{Method}

One of the artefacts of the MIRACLE project \cite{parker2019final} was the
Social Simulation REplication Interface or SSREPI. This is the schema shown in
\ref{fig:schema}.  This schema has evolved since it original conception.  The
newest version of this document may be found here \cite{polhill2022miracle}. This is a database
schema derived from the Dublin Core \cite{weibel2000dublin}, the standard XML
datatypes \cite{biron2004xml} and the PROV-O ontologies \cite{missier2013w3c}.
The schema is designed with agnosticisim towards the underlying database
technology and as such has been implmented both in PostgresSQL
\cite{stonebraker1991postgres} and Sqlite3 \cite{sqliteorg2023syntax}.

One of the aims was to leave as much of the code, i.e. the original experiment
as unchanged as possible. That is we wanted to change the framework code rather
than the core code. The core code in the example above being the executable,
precompiled in Objective-C \cite{objectivec}. Additionally the support code, i.e. the
code that prepared the data, a Perl \cite{perl563} script and the code that did the
post analysis, a mixture of Perl and R \cite{R422}.  This meant we developed the
database access in Python \cite{python368}, wrote wrapper scripts in Bash \cite{bash4420}. This
differs in approach to FAIR data piperline \cite{mitchell2022fair} in which the provenance framework is
embedded in the code that constitutes the body of the experiment as it were.

With the tool in the current state this involved labouriously going through the
code and inserting the some of the following Bash-based commands, shown in
table \ref{tab:commands}. These commands have a fairly complex parameter set as
it stands, and no way of checking whether the code worked other than running
it, and this led to one of the immediate problems with this method.

\begin{figure}
\includegraphics[width=\textwidth]{img/schema.jpeg}
\caption{SSREPI Schema} \label{fig:schema}
\end{figure}

There are two important dimensions of distinction to this schema. First is
fine- versus coarse- grained metadata. Second is provenance versus
workflow. Coarse-grained metadata describes how particular files come (or
came) into being, or were (or could be) used to bring other files into
being. Fine-grained metadata describes specific values recorded in social
simulation outputs. To make the distinction concrete, suppose a simulation
produces a CSV file. The data within the CSV file are covered by
fine-grained metadata, whilst the fact that the simulation produces the
CSV file is coarse-grained. Turning to the other dimension, provenance
metadata describes what actually happens (run W of simulation X produced
output file Y), whilst workflow metadata describes what could happen
(simulation X produces an output file of type Z). The distinctions are
summarised in \ref{fig:finegrain} .

\begin{figure}
\includegraphics[width=\textwidth]{img/fine-grain-vs-coarse-grain.jpeg}
\caption{Fine grain vs coarse grain provenance} \label{fig:finegrain}
\end{figure}

For the purposes of this short paper we are concentrating on a
demonstration of recording provenance. Indeed we use the example,
mentioned in the introduction of \cite{polhill2022miracle}, and modified the original Bash \cite{bash4420} scripts to include the what we denote as \textit{primitives} . So an example of the code

In order to record provenance we 
We coded the following primitives from
the SSREPI interface definition \cite{polhill2022miracle}. These are implemented in a mixture of Python and Bash.

\begin{table}
\caption{Table captions should be placed above the
tables.}\label{tab:commands}
    \begin{tabularx}{\textwidth}{|l|p{2cm}|p{6cm}|}
\hline
Primitive & Type & Purpose \\
\hline
        SSREPI\_require\_minimum & Metadata & Lower bound on software hardware required \\ SSREPI\_require\_exact & Metadata & Exact bound on software hardware required \\
        SSREPI\_application & Provenance \& Metadata & specifies some executable \\
        SSREPI\_me & Provenance \& Metadata & Determines executable being run or returns a proper reference to the executable being run. \\
        SSREPI\_run & Provenance \& metadata & Blocking invocation of an executable which will allow the specification of all inputs, outputs and arguments. Creates run-time provenance information as well \\
        SSREPI\_batch & Provenance \& metadata & Non blocking invocation of an executable which will allow the specification of all inputs, outputs and arguments. Creates run-time provenance information as well \\
        SSREPI\_argument & Provenance & An argument type to an exectuable \\
        SSREPI\_output & Provenance & An output type from an executable \\
        SSREPI\_input & Provenance & An input type for an exectuable \\
        SSREPI\_hutton\_person & Metadata & Uses our institutions databases to populate metadata for a given individual \\
        SSREPI\_person & Metdata & Provide metadata for a particular actor within this system\\
        SSREPI\_project & Metadata & Specifies a project which contains all studies \\
        SSREPI\_study & Metadata & A set of experiments makes up a single study \\
        SSREPI\_set & Metdata & Sets the default licence and other metadata \\
        SSREPI\_involvement & Metadata & Links personnel to a study \\
        SSREPI\_paper & Metadata & A paper associated with this study \\
        SSREPI\_make\_tag & Metadata & Used for building a folksonomy \\
        SSREPI\_tag & Metadata & Used to tag any entity with a folksonomy tag \\
        SSREPI\_contributor & Metadata & A  person with some kind of relation to an executable or script. \\

        SSREPI\_statistical\_method & Metadata & A statistical method is an approach to computing some statistics. It may be implemented in or as part of an application. A statistical method generates one or more statistical variables as its results, and may use the results of another statistical method in its computation. For example, computing the standard deviation of some data uses the mean of those data. \\
        SSREPI\_visualisation & Metadata & A visualisation is the process of creating an image to depict one or more (typically more than one) values. The results of a visualisation appear in a container.\\
        SSREPI\_statistics & Metadata & Statistics are activities that compute  and populate the values of statistical rvariables. They operate on raw data that are retrieved from the values using a query. To replicate a set of statistics, the query can be rerun, selecting values that are pointed to by containers entries \\
        SSREPI\_visualisation\_method & Metadata & This describes methods for generating visualisations, which then may appear in the content of a container produced by a process running an application that implements it. \\

        SSREPI\_implements & Metadata & Links a statistical or visualisation method to an application \\
        SSREPI\_parameter & Metadata & A Parameter is the name of a parameter taken by a statistical or visualisation method, used to configure the way it behaves. \\
        SSREPI\_statistical\_variable & Metadata &  A name for (one of) the result(s) of a statistical method. \\
        SSREPI\_visualisation\_variable & Provenance \& Metadata & Declares a named variable of interest \\
        SSREPI\_variable & Metadata & Names a variable of interest \\
        SSREPI\_statistical\_variable\_value & Provenance \& Metdata & Sets an actual value for a named statistical variable \\
        SSREPI\_value & Provenance & Sets an actual value. This can be connected to any metadata value such         \\ 
        SSREPI\_content & Metadata & Links a kind of output/input/argument to a variable  \\
        SSREPI\_person\_makes\_assumption & Metadata & links a person to an assumption \\
\hline
\end{tabularx}
\end{table}

Broadly speaking SSREPI\_application, SSREPI\_run, SSREPI\_batch, SSREPI\_input,
SSREPI\_output and SSREPI\_argument are responsible for recording coarse grain
provenance. SSREPI\_value, SSREPI\_visualisation\_variable\_value, SSREPI\_statistical\_variable\_value, SSREPI\_run and SSREPI\_batch record fine-grain provenance. The remaining
primitives are largely about recording metadata.

Elaboration of all this may be found in the man pages and design documents in the public repository at git@github.com:DougSalt/ABM-metadata.git

\begin{figure}
\includegraphics[width=\textwidth]{img/workflow.pdf}
\caption{The workflow sub-graph} \label{fig:workflow.pdf}
\end{figure}

\section{Results}

As mentioned the method undertaken in this paper involved inserting the correct
Bash-based commands, listed in table \ref{tab:commands}. These commands
have a fairly complex parameter set as it stands, and no way of checking
whether the code worked other than running it, and this led to one of the
immediate problems with this method. The only way to see if the code was
working was to run it, a suck it and see methodology. This was extremely time consuming. We took the obvious approach and ran small subsets of
the ~20,000 runs or so, to see if we could get the framework to work and then scaled it up to the full run, but invariably we ran into new problems the greater the scale. It therefore took a substantial time to even get this first replication to run. It took several months of coding effort to get this correct.

Obviously with complicated diagrams result from the 20,000 runs. The challenge becomes how to visualise these.
There are several graphs produced so far, these being

We show the workflow graph in fig. \ref{fig:workflow} and a section of the
proveance graph in fig. \ref{fig:sub-provenance}

So what use is this provenence meta data? Beyond replicability

\begin{figure}
\includegraphics[width=\textwidth]{img/subsection-of-provenance.png}
\caption{A small sub-section of the proevenance graph} \label{fig:sub-provenance}
\end{figure}

Those with a sharp eye may have noticed the versions of Perl, Python and R were not around when the original experiments were run and obviously this has implications.

\section{Discussion}

What we have learnt from this?  As noted the process of running the experiment with the new provenance and metadata commands was a complete pfaff. It involved the repeated running of the experiment with a larger and larger array of inputs until we matched the size of the original experiment.  There were codes in the both the Python and bash code (calling the Python scripts)

This has taken me almost a year to get it running. Repeating an experiment in a completely automated fashion turns out tbe a really hard task. The plan is to adapt this provenance framework to other model running language frameworks, in particular R and Python.

We demonstrate this by tracing a series of bad data across the diagrams.

The proveance should be a directed acyclic graph. There may well be loops in the schema as it stands. The authors have done their best to normalise it to uncover such cycles, but in a schema this complex they may well still exist.

Eventually we should like to take several databases run some machine learning across them to see if there are any commonalities in experiment set up and post processing.

Additionally we want to produce modules in other languages as well, such as R, python and Julia. Especially the latter.

One of the nice things was that I was able to parallelise the post processing job simply.

The underlying database is probably wrong. We should be using a graphing language such as Gremlin

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{global.bib}
% \bibliography[keyword={github.com/DougSalt/SSC2023}]{global.bib}
%
\end{document}